---
title: "Homework 2"
author: "Elliot Kim"
date: "2024-10-02"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r library}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(readxl)
library(tidyr)
```

## Problem 1

Import subway dataset and clean variables
```{r}
nyc_subway = read.csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv", na = c(".", "NA", "")) |> 
  janitor::clean_names() |>
  select(line:entry, vending, ada) |>
  mutate(entry = ifelse(entry =="YES", TRUE, FALSE))
```

The dataset above is called nyc_subway, which contains the following variables about nyc subways: the lines, station names, station latitude/longitude, routes(1-11), entrance type, entry, vending, and ADA compliance. To clean this dataset, I first read the csv file and standardized the column names. I then selected the variables listed above using the select function. Finally, I modified the entry variable by converting it from a character to a logical variable using the ifelse function. This dataset has `r nrow(nyc_subway)` rows and `r ncol(nyc_subway)` columns. The data is tidy as each variable has a distinct column and each observation occupies a distinct row. 

Dataset questions and answers
```{r}
distinct(nyc_subway, line, station_name) |>
  nrow()
```

There are 465 distinct stations in this dataset.

```{r}
filter(nyc_subway, ada == TRUE) |>
  distinct(line, station_name) |>
  nrow()
```

84 stations are ADA compliant.

```{r}
filter(nyc_subway, vending == "NO", entry == "TRUE") |>
  summarise(proportion = n()/nrow(nyc_subway))
  
```

The proportion of station entrances/exits without vending allow entrances is 3.69%

##reformat data
```{r}
nyc_subway2 = nyc_subway |>
  mutate(across(route1:route11, as.character)) |>
  pivot_longer(cols = route1:route11, names_to = "route_number", values_to = "route")

nyc_subway2 |>
  filter(route == "A") |>
  distinct(line, station_name) |>
  count()
  
nyc_subway2 |>
  filter(route == "A", ada == TRUE) |>
  distinct(line, station_name) |>
  count()
```

There are 60 dustubct stations that serve the A train. Of the stations that serve the A train, 17 are ADA compliant. 

## Problem 2
```{r}
trash_wheel = read_excel("./data/202409 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel")

trash_wheel |>
  janitor::clean_names() |>
  select(dumpster:homes_powered) |>
  drop_na(dumpster)
```

